---
title: "Week4Project1DATA607"
author: "Banu Boopalan"
date: "9/22/2019"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
  ioslides_presentation: default
  pdf_document: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Install Downloader
library(downloader)
library(stringr)
#Install Reshape pacakge 
#Start building data needed to calculate final average prechess rating
library(reshape)
library(tibble)
library(dplyr)
library(knitr)
#Install KableExtra package and use that
library(kableExtra)
```

## Section1:PROJECT 1: Read Chess Rating dataset and create a file to write to CSV. First Read file from my GITHUB url.

Read dataset from url on github

```{r}
myurl <- "https://raw.githubusercontent.com/BanuB/Week4AssignmentDATA607/master/tournamentinfo.txt"
x = read.csv(file=myurl)
filename <- "tournamenttxtfile.txt"

downloader::download(myurl, filename)
df <- readLines(filename)

myconn <- file(filename,open="r")
myfiledata <- readLines(filename,warn = FALSE)
close(myconn)

head(myfiledata)
ncol(myfiledata)
```

## Section2:Subset
Final 2 Subsets to extract and separate and join back to get only 1 row for each player

```{r}
#Final 2 Subsets to extract and separate and join back to get only 1 row for each player
#------------------------------------------------------------------------

subset1 <- myfiledata[str_detect(myfiledata,"([0-9]{1}[|])")]
subset1


subset2 <- myfiledata[str_detect(myfiledata,"([:]{1})")]
subset2

```

## Section3:CBIND
Start with this dataset which is a cbind of rows for each player in 1 row

```{r}
#--------------------------------------------------------------------------
#Start with this dataset which is a cbind of rows for each player in 1 row

df1 <- data.frame(subset1)
df2 <- data.frame(subset2)
df3 <- cbind (df1, df2)
str(df3)
df3
```

## Section4:String Locate
String locate to locate positions to extract our subtring to build the final players dataframe. Using only 1st row since we only need that to determine positions for certain field and rest of the rows will follow the same positions to extract from

```{r}

str_locate_all(df3$subset1[1],"\\b[:alnum:]+\\b")
str_locate_all(df3$subset2[1],"\\b[:alnum:]+\\b")

```

## Section5:First Main chess dataframe to build final columns to display. Convert player_num and player_prerating, total_points to decimal


```{r}
#------------------------------------------------------------------------------------------------------------------------
#First Main chess dataframe to build final columns to display. Convert player_num and player_prerating, total_points to decimal

startnew <- data.frame(player_num = type.convert((substr(df3$subset1,4,5)),na.strings = "NA", as.is = FALSE, dec = "."), 
                       player_name = paste(c(str_extract(df3$subset1,"\\b([A-Z]+)\\s*([A-Z]+)\\s*([A-Z]+)\\s*([A-Z]+)\\b"))), 
                       State = substr(df3$subset2,4,5),
                       #                       total_points = paste(c(str_extract(df3$subset1,"(\\d\\.)[0-9]"))),
                       total_points = type.convert(str_extract(df3$subset1,"(\\d\\.)[0-9]"), na.strings = "NA", as.is = FALSE, dec= "."),
                       player_prerating = type.convert((substr(df3$subset2,23,26)),na.strings = "NA", as.is = FALSE, dec = ".")) 

startnew
str(startnew)


```

## Section6:Use Melt of melt.list of opponentnumber to get player number with the corresponding opponent number


```{r}
#------------------------------------------------------------------------------------------------------------------------
#melt list with opponentnumber to get player with the opponentnumber and put that in a dataframe
opponentnum <-  type.convert(str_extract_all(df3$subset1,"\\b\\s+([0-9]+)\\b"),na.strings = "NA", as.is = FALSE, dec = ".")
opponentnum
str(opponentnum)
df <- melt(opponentnum)


```

## Section7:View melted dataset and reassign column names for merging later


```{r}
#--------------------------------------------------------------------------------------------------------------------------------
#view melted dataframe and reassign column names to a different name
df <- setNames(df,c("oppo_num","player_num"))
str(df)
df

```

## Section8:Merge melted dataframes by player_number to get player number along with the opponents played


```{r}
#--------------------------------------------------------------------------------------------------------------------------------
#Merge melted dataframe with dataframe below by player_num to get rows of player with opponent number. Here any games with 
#missing opponent number is not kept. For instance the Melt only reads in the 5 games that were played for player #16 since the other 2 games don't have a opponent # listed in the 
#string that was read in for this player
df2 <- data.frame(player_num = type.convert(substr(df3$subset1,4,5),na.strings = "NA", as.is = FALSE, dec = "."), 
                  player_name = paste(c(str_extract(df3$subset1,"\\b([A-Z]+)\\s*([A-Z]+)\\s*([A-Z]+)\\s*([A-Z]+)\\b"))),
                  player_prerating = type.convert((substr(df3$subset2,23,26)),na.strings = "NA", as.is = FALSE, dec = "."))
str(df2)
df2

df4 <- merge(df2,df,by ="player_num")
head(df4)
str(df4)

```


## Section9:Merge again and order final df8 dataset by player number for group_by and aggregating rows


```{r}
#--------------------------------------------------------------------------------------------------------------------------------
#new dataset to reassign just player prerating and player number as opponent prerating and opponent playernumber
df6 <- data.frame(oppo_num = df2$player_num, oppo_prerating = df2$player_prerating)
df6 

#--------------------------------------------------------------------------------------------------------------------------------
#merge to get oppo prerating
df7 <- merge(df4, df6, by = "oppo_num")
df7
str(df7)

#--------------------------------------------------------------------------------------------------------------------------------
#order the df8 dataset by player# so we can group to find average oppponent prechess rating
df8 <- df7[
  order( df7[,2] ),
  ]
df8

```

## Section10:Aggregate rows and group by player_num to get average opponent pre chess rating

```{r}
chessrating_df <- df8 %>% group_by(player_num) %>% summarize(avg_prechess_rating = mean(oppo_prerating))
str(chessrating_df)
chessrating_df

```

## Section11:Final merge to get all columns  and write out kable table

```{r}
#merge back to main dataset to get the finaldataset by player_number with all the fields
finaldf <- merge(startnew, chessrating_df, by = "player_num")
finaldf
finaldf %>% kable() %>% kable_styling()
```

## Section12:Write final CSV files

```{r}
#Write file to a CSV file
write.csv(finaldf, "Banufinalchessratingdataset.csv", row.names = FALSE)
write.csv(df8, "Banuplayerrowswithopponentrating.csv", row.names = FALSE)
```

